---
title: "Next Generation Attibution"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<div style="text-align: right"><i><h5>Dimitris Psaradellis, Pavol Gurican, Elias Ron, Rachael Maltiel-Swenson</h5></i></div>

```{r libraries, echo=FALSE, message=FALSE}
setwd("~/projects/next_gen_attribution/src/analysis/R/pg")

source("~/R_connector/libraries.R")
#mozart <- connToEDW(dataBase = "Mozart", connType = 'normal', user = 'pgurican', pass=trimws(read_file("~/R_connector/p.txt")))
hopper <- connToEDW(dataBase = "Hopper", connType = 'normal', user = 'pgurican', pass=trimws(read_file("~/R_connector/p.txt")))

library(htmlwidgets)
library(highcharter)
library(plotly)

cnts <- dbGetQuery(conn = hopper, "Select channel_name, day_diff, sum(cnt) as cnt from P_TRAFFIC_T.DP_TESTING_MOVER where event_type_id in (1, 5, 7) group by 1, 2")

excluded <- cnts %>% group_by(channel_name) %>% arrange(day_diff) %>% mutate(cumm_cnt=cumsum(cnt)) %>% 
  select(channel_name, day_diff, cumm_cnt) %>% filter(day_diff==61) %>% arrange(cumm_cnt) %>% select(-day_diff) %>% 
  filter(cumm_cnt < 1000000) %>%  select(channel_name) %>% unlist()

excluded <- c(excluded, "NULL", "isNA")

```

## Marketing Events

### All Events

We start with plotting the relative proportions of all marketing events (clicks, impressions, opens, ...) happening in all channels over a given time window.

```{r marketing_events, echo=FALSE, message=FALSE, warning=FALSE}
mktng_events <- dbGetQuery(conn = hopper, "select * from P_attribution_T.pg_agg_mktng_evts_v4;")
mktng_events <- mktng_events %>% select(-dt)
mktng_events$event_dt <- as.Date(mktng_events$event_dt)
mktng_events$channel_id <- as.factor(mktng_events$channel_id)
mktng_events$event_type_id <- as.factor(mktng_events$event_type_id)
mktng_events$experience_level2 <- as.factor(mktng_events$experience_level2)
mktng_events$device_level2 <- as.factor(mktng_events$device_level2)

channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events <- left_join(mktng_events, channels_lookup)

events_by_date <- mktng_events %>% 
                                    select(-event_type_id) %>% 
                                    group_by(event_dt, channel_name) %>% 
                                    summarise(no_events = sum(no_events))

ds <- lapply(levels(events_by_date$channel_name),
             function(x){
                 list(data = (events_by_date %>% 
                                      filter(channel_name == x) %>% 
                                      arrange(event_dt))$no_events,
                      name = x,
                      visibility = !(x %in% excluded))
             })

hc <- highchart() %>% 
  hc_chart(type = "area") %>% 
  hc_title(text = "Proportion of marketing events per channel") %>% 
  hc_subtitle(text = "Next Generation Attribution") %>% 
  hc_xAxis(categories = seq(from = as.Date("2017-05-01"),
                            to = as.Date("2017-07-01"),
                            by = 1),
                       #(unique(events_by_date$event_dt)),
           tickmarkPlacement = "on",
           title = list(text = "Date of Marketing Event")) %>% 
  hc_yAxis(title = list(text = "Percent")) %>% 
  hc_tooltip(pointFormat = "<br/><span style=\"color:{series.color}\">{series.name}</span>:
             <b>{point.percentage:.1f}%</b> ({point.y:,.0f} events)",
              crosshairs = TRUE, 
          borderWidth = 3, 
          #sort = TRUE, 
          table = TRUE) %>% 
  hc_plotOptions(area = list(
     stacking = "percent",
     lineColor = "#ffffff",
     lineWidth = 1,
     marker = list(
       lineWidth = 1,
       lineColor = "#ffffff"
       ))
     ) %>%
    hc_add_series_list(ds) %>%
  hc_size(height = 700) %>% 
  hc_exporting(
    enabled = TRUE
  )
```

```{r plot_mktng_events, echo=FALSE, warning=FALSE}
hc
```

The second plot shows the same information, but in absolute scale.

```{r plot_absolute_mktng_events, echo=FALSE, warning=FALSE, message=FALSE}
hc <- highchart() %>% 
  hc_chart(type = "area") %>% 
  hc_title(text = "Absolute numbers of marketing events per channel") %>% 
  hc_subtitle(text = "Next Generation Attribution") %>% 
  hc_xAxis(categories = as.Date(seq(from = as.Date("2017-05-01"),
                            to = as.Date("2017-07-01"),
                            by = 1)),
                       #(unique(events_by_date$event_dt)),
           tickmarkPlacement = "on",
           title = list(text = "Date of Marketing Event")) %>% 
  hc_yAxis(title = list(text = "Number of Events")) %>% 
  hc_tooltip(pointFormat = "<br/><span style=\"color:{series.color}\">{series.name}</span>:
             <b>{point.percentage:.1f}%</b> ({point.y:,.0f} events)",
             shared = TRUE,
              crosshairs = TRUE, 
          borderWidth = 3, 
          #sort = TRUE, 
          table = TRUE) %>% 
  hc_plotOptions(area = list(
     stacking = "normal",
     lineColor = "#ffffff",
     lineWidth = 1,
     marker = list(
       lineWidth = 1,
       lineColor = "#ffffff"
       ))
     ) %>%
    hc_add_series_list(ds)  %>%
    hc_size(height = 700) %>% 
  hc_exporting(
    enabled = TRUE
  )

hc
```

### Clicks only

```{r marketing_events_clicks, echo=FALSE, message=FALSE, warning=FALSE}
mktng_events <- dbGetQuery(conn = hopper, "select * from P_attribution_T.pg_agg_mktng_evts_v4 where event_type_id in (1,5,7);")
mktng_events <- mktng_events %>% select(-dt)
mktng_events$event_dt <- as.Date(mktng_events$event_dt)
mktng_events$channel_id <- as.factor(mktng_events$channel_id)
mktng_events$event_type_id <- as.factor(mktng_events$event_type_id)
mktng_events$experience_level2 <- as.factor(mktng_events$experience_level2)
mktng_events$device_level2 <- as.factor(mktng_events$device_level2)

channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events <- left_join(mktng_events, channels_lookup)

events_by_date <- mktng_events %>% 
                                    select(-event_type_id) %>% 
                                    group_by(event_dt, channel_name) %>% 
                                    summarise(no_events = sum(no_events))

ds <- lapply(levels(events_by_date$channel_name),
             function(x){
                 list(data = (events_by_date %>% 
                                      filter(channel_name == x) %>% 
                                      arrange(event_dt))$no_events,
                      name = x)
             })

hc <- highchart() %>% 
  hc_chart(type = "area") %>% 
  hc_title(text = "Proportion of clicks per channel") %>% 
  hc_subtitle(text = "Next Generation Attribution") %>% 
  hc_xAxis(categories = seq(from = as.Date("2017-05-01"),
                            to = as.Date("2017-07-01"),
                            by = 1),
                       #(unique(events_by_date$event_dt)),
           tickmarkPlacement = "on",
           title = list(text = "Date of Click")) %>% 
  hc_yAxis(title = list(text = "Percent")) %>% 
  hc_tooltip(pointFormat = "<br/><span style=\"color:{series.color}\">{series.name}</span>:
             <b>{point.percentage:.1f}%</b> ({point.y:,.0f} events)",
              crosshairs = TRUE, 
          borderWidth = 3, 
          #sort = TRUE, 
          table = TRUE) %>% 
  hc_plotOptions(area = list(
     stacking = "percent",
     lineColor = "#ffffff",
     lineWidth = 1,
     marker = list(
       lineWidth = 1,
       lineColor = "#ffffff"
       ))
     ) %>%
    hc_add_series_list(ds) %>%
  hc_size(height = 700) %>% 
  hc_exporting(
    enabled = TRUE
  )

hc
```

## Marketing events with correlations to transactions

### Relative importances of channels with or without transactions

``` {r plot_proportions, echo=FALSE, message=FALSE, warning=FALSE}
mktng_events <- dbGetQuery(conn = hopper, "select * from P_attribution_T.pg_agg_mktng_evts_v3;")
mktng_events <- mktng_events %>% select(-dt)

mktng_events$event_dt <- as.Date(mktng_events$event_dt)
mktng_events$channel_id <- as.factor(mktng_events$channel_id)
mktng_events$event_type_id <- as.factor(mktng_events$event_type_id)
mktng_events$experience_level2 <- as.factor(mktng_events$experience_level2)
mktng_events$device_level2 <- as.factor(mktng_events$device_level2)

channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events <- left_join(mktng_events, channels_lookup)

mktng_events_prop <- mktng_events %>% 
        select(-event_type_id, channel_id) %>% 
        group_by(event_dt, channel_name) %>% 
        summarise(daily_events_channel = sum(no_events)) %>%
        ungroup(channel_name) %>%
        group_by(event_dt) %>%
        mutate(daily_events = sum(daily_events_channel)) %>%
        mutate(prop_evts = daily_events_channel/daily_events*100) %>%
        arrange(event_dt, channel_name)


mktng_events_07_01 <- dbGetQuery(conn = hopper, "select * from P_attribution_T.mktng_txns_v4_bbowa_07_01;")

mktng_events_07_01 <- mktng_events_07_01 %>% select(-dt)
mktng_events_07_01$event_dt <- as.Date(mktng_events_07_01$event_dt)
mktng_events_07_01$channel_id <- as.factor(mktng_events_07_01$channel_id)
mktng_events_07_01$event_type_id <- as.factor(mktng_events_07_01$event_type_id)
mktng_events_07_01$experience_level2 <- as.factor(mktng_events_07_01$experience_level2)
mktng_events_07_01$device_level2 <- as.factor(mktng_events_07_01$device_level2)
channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events_07_01 <- left_join(mktng_events_07_01, channels_lookup)

mktng_events_txns_daily_prop_07_01 <- mktng_events_07_01 %>% 
        select(-event_type_id, channel_id) %>% 
        group_by(event_dt, channel_name) %>% 
        summarise(daily_events_channel = sum(no_events)) %>%
        ungroup(channel_name) %>%
        group_by(event_dt) %>%
        mutate(daily_events = sum(daily_events_channel)) %>%
        mutate(prop_evts_txns = daily_events_channel/daily_events*100) %>%
        arrange(event_dt, channel_name)

proportions <- left_join(mktng_events_txns_daily_prop_07_01 %>%
                                select(event_dt,
                                       channel_name,
                                       prop_evts_txns),
                         mktng_events_prop %>% select(event_dt,
                                                      channel_name,
                                                      prop_evts),
                         by = c('event_dt', 'channel_name'))

hc <- hchart(proportions,
             type = "line",
             hcaes(x = event_dt,
                   y = prop_evts_txns/prop_evts,
                   group = channel_name)) %>%
        hc_xAxis(title = list(text = "Date of the Marketing Event")) %>%
        hc_yAxis(title = list(text = "Ratio")) %>%
        hc_title(text = "Relative proportion of traffic tied to a BBOWAC event on 07/01 attributed to channel with respect to overall marketing-driven traffic.") %>%
        hc_tooltip( pointFormat = "<br/><span style=\"color:{series.color}\">{series.name}</span>:
            <b>{point.y:,.2f}x</b>",
             #shared = TRUE
          crosshairs = TRUE, 
          borderWidth = 3, 
          sort = TRUE, 
          table = TRUE) %>% 
        hc_size(height = 700) %>% 
  hc_exporting(
    enabled = TRUE
  )

hc


```

``` {r plot_proportions2, echo=FALSE, message=FALSE, warning=FALSE}
mktng_events <- dbGetQuery(conn = hopper, "select * from P_attribution_T.pg_agg_mktng_evts_v3;")
mktng_events <- mktng_events %>% select(-dt)

mktng_events$event_dt <- as.Date(mktng_events$event_dt)
mktng_events$channel_id <- as.factor(mktng_events$channel_id)
mktng_events$event_type_id <- as.factor(mktng_events$event_type_id)
mktng_events$experience_level2 <- as.factor(mktng_events$experience_level2)
mktng_events$device_level2 <- as.factor(mktng_events$device_level2)

channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events <- left_join(mktng_events, channels_lookup)

mktng_events_prop <- mktng_events %>% 
        select(-event_type_id, channel_id) %>% 
        group_by(event_dt, channel_name) %>% 
        summarise(daily_events_channel = sum(no_events)) %>%
        ungroup(channel_name) %>%
        group_by(event_dt) %>%
        mutate(daily_events = sum(daily_events_channel)) %>%
        mutate(prop_evts = daily_events_channel/daily_events*100) %>%
        arrange(event_dt, channel_name)


mktng_events_06_22 <- dbGetQuery(conn = hopper, "select * from P_attribution_T.mktng_txns_v4_bbowa_06_22;")

mktng_events_06_22 <- mktng_events_06_22 %>% select(-dt)
mktng_events_06_22$event_dt <- as.Date(mktng_events_06_22$event_dt)
mktng_events_06_22$channel_id <- as.factor(mktng_events_06_22$channel_id)
mktng_events_06_22$event_type_id <- as.factor(mktng_events_06_22$event_type_id)
mktng_events_06_22$experience_level2 <- as.factor(mktng_events_06_22$experience_level2)
mktng_events_06_22$device_level2 <- as.factor(mktng_events_06_22$device_level2)
channels_lookup <- read.csv(file = "channels_lookup.csv")
channels_lookup$channel_id <- as.factor(channels_lookup$channel_id)
channels_lookup$channel_name <- as.factor(channels_lookup$channel_name)

mktng_events_06_22 <- left_join(mktng_events_06_22, channels_lookup)

mktng_events_txns_daily_prop_06_22 <- mktng_events_06_22 %>% 
        select(-event_type_id, channel_id) %>% 
        group_by(event_dt, channel_name) %>% 
        summarise(daily_events_channel = sum(no_events)) %>%
        ungroup(channel_name) %>%
        group_by(event_dt) %>%
        mutate(daily_events = sum(daily_events_channel)) %>%
        mutate(prop_evts_txns = daily_events_channel/daily_events*100) %>%
        arrange(event_dt, channel_name)

proportions2 <- left_join(mktng_events_txns_daily_prop_06_22 %>%
                                select(event_dt,
                                       channel_name,
                                       prop_evts_txns),
                         mktng_events_prop %>% select(event_dt,
                                                      channel_name,
                                                      prop_evts),
                         by = c('event_dt', 'channel_name'))

hc <- hchart(proportions2,
             type = "line",
             hcaes(x = event_dt,
                   y = prop_evts_txns/prop_evts,
                   group = channel_name)) %>%
        hc_xAxis(title = list(text = "Date of the Marketing Event")) %>%
        hc_yAxis(title = list(text = "Ratio")) %>%
        hc_title(text = "Relative proportion of traffic tied to a BBOWAC event on 06/22 attributed to channel with respect to overall marketing-driven traffic.") %>%
        hc_tooltip( pointFormat = "<br/><span style=\"color:{series.color}\">{series.name}</span>:
            <b>{point.y:,.2f}x</b>",
             #shared = TRUE
          crosshairs = TRUE, 
          borderWidth = 3, 
          sort = TRUE, 
          table = TRUE) %>% 
        hc_size(height = 700) %>% 
  hc_exporting(
    enabled = TRUE
  )

hc


```

## Correlated Transactions

The average number of clicks for correlated transaction is **3.31** with a 1-day window, **4.05** for a 7-day window and **6.26** for a window of one month. 

Detailed breakdown according to the marketing channel can be found below.

``` {r correlated_txns, echo=FALSE, message=FALSE, warning=FALSE}
require(DT)
averageClicks <- read_csv("~/projects/next_gen_attribution/src/analysis/R/pg/averageClicks.csv")
names(averageClicks)[1] <- "channel_name"

datatable(
    averageClicks,
    rownames = FALSE,
    # caption = 'Table 1: This is a simple caption for the table.',
    extensions = 'Buttons',
    options = list(
      columnDefs = list(list(autoWidth = TRUE, className = 'dt-left', targets = c(0))),
      pageLength = 50,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv')
    )) %>%
  formatRound(names(averageClicks)[-1], 2)

```

### Lookback period overview

From a marketing channel viewpoint

``` {r txns_by_channel, echo=FALSE, message=FALSE, warning=FALSE}


# cnts2 <- cnts
cnts %<>% group_by(channel_name) %>% mutate(prcnt=round(cnt/sum(cnt)*100, 1)) %>% ungroup()
cnts$channel_name[is.na(cnts$channel_name)] <- "isNA"


cnts3 <- cnts %>% group_by(channel_name) %>% arrange(day_diff) %>% mutate(cumm_cnt=cumsum(cnt)) %>% 
  select(channel_name, day_diff, cumm_cnt) %>% spread(channel_name, cumm_cnt)



hchart(cnts %>% arrange(channel_name, day_diff) %>% group_by(channel_name) %>% 
         mutate(cumm_cnt=cumsum(cnt), cumm_prcnt=cumsum(prcnt)),
       "line", 
       hcaes(x = day_diff, 
             y = cumm_prcnt, 
             group = channel_name),
       visible = !(unique((cnts %>% arrange(channel_name))$channel_name) %in% excluded)) %>% 
  hc_xAxis(title = list(text = "Days from Marketing event to first BBOWAC event")) %>% 
  hc_yAxis(title = list(text = "Cummulative % Correlated Transactions"), max = 100) %>% 
  hc_exporting(enabled = TRUE)



```


Examining the number of correlated transactions by lookback window

```{r lookback_window, echo=FALSE, message=FALSE, warning=FALSE}
hc <- highchart() %>% hc_chart(type = "column") %>% hc_xAxis(categories = " ")
for(x in names(cnts3)[-1]){
  hc <- hc %>% hc_add_series(name = x, visible = ifelse(x %in% excluded, FALSE, TRUE),
                             data = list(list(sequence =    setNames(unlist(cnts3[x]), NULL)    )))
}
hc <- hc %>% hc_motion(enabled = TRUE, labels = 0:61, 
                       series = 1:(length(names(cnts3)[-1])-1))
hc
```

&nbsp;

&nbsp;

&nbsp;


### Average number of clicks by channel and FM segment

Looking at the average number of clicks by channel and FM segment

``` {r avg_clicks_per_FM_segment, echo=FALSE, message=FALSE, warning=FALSE}
averageClicksBySegment <- read_csv("~/projects/next_gen_attribution/src/analysis/R/pg/bySegments.csv")

datatable(
     averageClicksBySegment,
    rownames = FALSE,
    # caption = 'Table 1: This is a simple caption for the table.',
    extensions = 'Buttons',
    options = list(
      columnDefs = list(list(autoWidth = TRUE, className = 'dt-left', targets = c(0))),
      pageLength = 50,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv')
    ))  %>%
  formatRound(names(averageClicksBySegment)[-1], 1)

```

### Transactions

``` {r transactions, echo=FALSE, message=FALSE, warning=FALSE}
averageClicksBySegment <- read_csv("~/projects/next_gen_attribution/src/analysis/R/pg/transactions.csv")

datatable(
     averageClicksBySegment,
    rownames = FALSE,
    # caption = 'Table 1: This is a simple caption for the table.',
    extensions = 'Buttons',
    options = list(
      columnDefs = list(list(autoWidth = TRUE, className = 'dt-left', targets = c(0))),
      pageLength = 50,
      dom = 'Bfrtip',
      buttons = c('copy', 'csv')
    ))  %>%
  formatRound(names(averageClicksBySegment)[2], 0)%>%
  formatRound(names(averageClicksBySegment)[3], 2)

